NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-01-30 main.tex
Updates from Overleaf

\author[1]{Anne Kiirikki}         %ORCID: ?
\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
\author[2]{Lara Bort}             %ORCID: ?
\author[3]{Pavel Buslaev}         %ORCID: ?
\author[4]{Fernando Favela}
\author[5]{Tiago Mendes Ferreira}
\author[6,7]{Patrick Fuchs}
\author[8,9]{Rebeca Garcia-Fandino}
\author[2]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[2]{Batuhan Kav}           %ORCID: ?
\author[2]{Milla Kurki}
\author[2]{Alexander Kuzmin}
\author[10,11]{Jesper Madsen}
\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[12]{Thomas Piggot}
\author[2]{Suman Samantray}
\author[2]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}

@@

\subsection{PCA analysis of equilibration of simulations}

@@

\section*{Acknowledgements}

%Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

\section*{Author contributions statement}

Must include all authors, identified by initials, for example:
A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

>>>>

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[14]{Lara Bort}             %ORCID: ?
\author[3]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[4]{Fernando Favela}
\author[5]{Tiago Mendes Ferreira}
\author[6,7]{Patrick F.J. Fuchs}
\author[8]{Rebeca Garcia-Fandino}
\author[2]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[2]{Batuhan Kav}           %ORCID: ?
\author[13]{Patrik Kula}
\author[2]{Milla Kurki}
\author[2]{Alexander Kuzmin}
\author[10,11]{Jesper Madsen}
\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[12]{Thomas Piggot}
\author[9]{{\'A}ngel Pi{\~n}eiro}
\author[2]{Suman Samantray}
\author[8,9,15]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[7]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[9]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[13]{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
\affil[15]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}

@@

\subsection{Principal Components Analysis of equilibration of simulations}
To estimate how well the provided trajectories are converged, the Principal Components Analysis (PCA) was used \cite{buslaev16,buslaev2020principal}. Following the PCALipids \cite{buslaev2020principal} protocol, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. Next, each lipid configuration was aligned to the average lipid structure. Afterwards, the PCA was applied on the cartesian coordinates of all heavy atoms of the lipid. Since it was shown before, that the motions along the first, major, principal component (PC) are the slowest ones \cite{buslaev16}, the equilibration of individual lipids was estimated based on the dynamics of the first PC. The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}. However, due to the linear dependence between autocorrelation decay times and distribution convergence times, faster calculation times and higher computational stability of autocorrelation decay times\cite{buslaev16, buslaev2020principal}, the distribution convergence timescales were calculated from autocorrelation decay times:
\begin{equation}
    \tau_{\text{convergence}} = 49*\tau_{\text{autocorrelation}}
\end{equation}

The empirical coefficient 49 was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields. All simulations, used for calculating the empirical coefficient, that relate convergence and autocorrelation, were at least 200 ns long. As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations as the most frequent in the Databank were used.

As a measure of equilibration quality for each trajectory, the ratio between the convergence time of first PC and the trajectory length was calculated for each lipid type individually. Sterols were excluded from the analysis. The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids databank repository. The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.

@@

\section*{Acknowledgements}
%Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
R.G.-F. thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09). 
AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00). 
O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.
\section*{Author contributions statement}
%Must include all authors, identified by initials, for example:
%A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
